Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-08-21 manuscript/manuscript.Rmd
updates ms with Britta's comments

Abstract
--------
Forecasting future states of populations has taken on new urgency as the rate of climate change increases. Traditional plant population models, typically built using detailed demographic observations from one study site, can address climate change impacts at one location, but are difficult to scale up to the landscape and regional scales relevant to management decisions. One proposal to overcome the limitations of traditional population models is to build models using population-level data that is much easier to collect over broad spatial scales than individual-level data. However, such models ignore the fact that climate drives population growth through its influence on individual performance. Here, we test the consequences of aggregating individual responses when forecasting climate change impacts on four perennial grass species in a semi-arid Montana grassland. We parameterized two population models, one based on individual-level data (survival, growth and recruitment) and one on population-level data from the same dataset (percent cover of quadrats), and compared their accuracy, precision, and sensitivity to climate. The individual level model was more accurate and precise than the aggregate level model when predicting out-of-sample observations. When comparing climate effects from both models, the population-level model tends to "miss" important climate effects from at least one vital rate for each species. Increasing the sample size at the population-level would not necessarily reduce forecast uncertainty; the way to reduce uncertainty is to capture unique climate dependence of individual vital rates. Our analysis indicates that there is no short cut to forecasting climate change impacts on plant populations — detailed demographic data is essential. Despite the superiority of the individual-level model, the forecasts it generated still were too uncertain to be useful for decision-makers. We need new methods to collect demographic data more efficiently across environmental gradients in space and time.

@@

Perhaps the greatest challenge for ecology in the 21st century is to forecast the impacts of environmental change [@Clark2001; @Petchey2015]. Forecasts require sophisticated modeling approaches that fully account for uncertainty and variability in both ecological process and model parameters [@Luo2011a]. This requires large amounts of data collected over broad spatio-temporal extents, and state-of-the-science modeling techniques cannot overcome data limitations. Population models are a case in point.

@@

The inability of most population models to address landscape-scale problems may explain why land managers and conservation planners have embraced species distribution models (SDMs) [see @Guisan2005 for a review]. SDMs typically rely on easy-to-collect presence/absence data [but see @Clark2014 for new methods] and remotely-sensed environmental covariates that allow researchers to model large spatial extents [e.g., @Maiorano2013]. Thus, it is relatively straightforward to parameterize and project SDMs over landscapes and regions, the scales at which many land-use decisions are made. However, the limitations of SDMs are well known [@Pearson2003; @Elith2009; @Araujo2012]. Ideally, researchers would provide managers with landscape-scale population models, combining the extent of SDMs with information about dynamics and species abundances.

@@

To date, most empirical population modelers have relied on individual-level data, with few attempts to exploit population-level data. An important exception was an effort by @Taylor2004 to model the population growth rate of an invasive species to identify the best strategies for invasion control. They used a "density-structured" model where the state variable is a discrete density state rather than a continuous density measure. Such models do not require individual level demographic data and can adequately describe population dynamics.

@@

Addressing climate change questions with models fit to population-level data is potentially problematic. It is individuals, not populations, that respond to climate variables [@Clark2012]. Ignoring this fact amounts to an "ecological fallacy", where inference about the individual relies on statistical inference on the group [@PIANTADOSI1988]. Population growth (or decline) is the outcome of demographic processes like survival, growth, and recruitment that occur at the level of individual plants. Climate can affect each demographic process in unique, potentially opossing, ways [@Dalgleish2011]. These unique climate responses may be difficult to resolve in statistical models based on population-level data where demographic processes are not identifiable. If population-level data cannot detect important impacts of climate drivers on populations, then population models built with such data will make poor forecasts.

Here, we ask whether statistical and population models based on aggregated, population-level data can predict climate change impacts as wells as models based on individual-level data. We used a unique demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. In both models, interannual variation is explained, in part, by climate covariates. We then performed simulations to quantify the sensitivities of species' cover to small perturbations in average precipitation and temperature. Individual level models were clearly better, but, unfortunately, still too uncertain to inform management decisions.

@@

An important, but unexpected, result of our analysis was the great uncertainty in forecasts, even for our best model. The typical approach in ecology is to use point estimates of model parameters to project populations forward according to the specified model, usually allowing for some variability around the determinstic process [e.g. @Battin2007; @Jenouvrier2009; @Adler2012]. If we follow tradition and calculate the mean response to climate perturbation with only model error and interannual variation included, the IPM and the QBM produce opposing forecasts for three of four species (Fig. D1). It would be tempting to interpret this inconsistency as further evidence for the superiority of the IPM. However, if we introduce parameter uncertainty, the forecasts are actually indistinguishable (Fig. 6), though the IPM projections are generally more precise (consistent with our cross-validation results). The real story is that both models produce highly uncertain forecasts. For all species, the 90% quantiles of predicted changes in population size overlapped zero; we cannot even predict whether a change in precipitation or temperature will cause populations to increase or decrease. \todo{Add in sentence on larger perturbations.}

@@

How might we improve on this state of affairs? First, forecasts could be improved by matching the spatial scale of predictor variables with the spatial scale of observations. One of the major limitations of the models we fit here is that the climate data are at a much larger scale than the individual level observations of plant size. Climate covariates only vary by year, with no spatial variability within years. Thus, even if we fit models to individual-level data, we are missing the key interaction point between weather and individual plants [@Clark2011b] because all observations share the same climate covariates. Demographic studies should be designed with at least plot-level measurements of climate related variables (e.g., soil moisture). Second, accurately detecting climate signals will take even longer time series. Recent theoretical work on detecting climate signals in noisy data suggests that even advanced approaches to parameter fitting require 20-25 year time series (Teller et al., in review). Third, ecologists need a stronger commitment to reporting uncertainty. Although most modeling studies explicitly consider model uncertainty, parameter uncertainty is often ignored. In some cases this is because the most convenient statistical methods make it difficult to propogate parameter uncertainty. Yet even Bayesian approaches that allow integration of model fitting and forecasting [@Hobbs2015] are not simple when using modeling approaches like integral projection models that separate the model fitting and simulation stages [@Rees2009]. However, as we have done here, it is still possible to include parameter uncertainty by drawing parameter values from MCMC iterations, taking care to draw all parameters from the same chain and iteration to account for their correlations. Only by being honest about our forecasts can we begin to produce better ones, and forecasts reported without parameter error are disingenuous. In many cases, ignoring parameter error is justifiable when investigating basic demographic mechanisms, but it is not justifiable when forecasting is the goal.

@@

Acknowledgments
---------------
This work was funded by the National Science Foundation through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040). We thank the original mappers of the permanent quadrats in Montana and the digitizers in the Adler lab, without whom this work would not have been possible. Informal conversations with Stephen Ellner, Giles Hooker, Robin Snyder, and a series of meetings between the Adler and Weecology labs at USU sharpened our thinking. Compute, storage and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged.

>>>>

Abstract
--------
Forecasting future states of populations has taken on new urgency as the rate of climate change increases. Traditional plant population models, typically built using detailed demographic observations from one study site, can address climate change impacts at one location, but are difficult to scale up to the landscape and regional scales relevant to management decisions. One proposal to overcome the limitations of traditional population models is to build models using population-level data that is much easier to collect over broad spatial scales than individual-level data. However, such models ignore the fact that climate drives population growth through its influence on individual performance. Here, we test the consequences of aggregating individual responses when forecasting climate change impacts on four perennial grass species in a semi-arid Montana grassland. We parameterized two population models, one based on individual-level data (survival, growth and recruitment) and one on population-level data from the same dataset (percent cover of quadrats), and compared their accuracy, precision, and sensitivity to climate. The individual level model was more accurate and precise than the aggregate level model when predicting out-of-sample observations. When comparing climate effects from both models, the population-level model tends to miss important climate effects from at least one vital rate for each species. Increasing the sample size at the population-level would not necessarily reduce forecast uncertainty; the way to reduce uncertainty is to capture unique climate dependence of individual vital rates. Our analysis indicates that there is no shortcut to forecasting climate change impacts on plant populations — detailed demographic data is essential. Despite the superiority of the individual-level model, the forecasts it generated still were too uncertain to be useful for decision-makers. We need new methods to collect demographic data more efficiently across environmental gradients in space and time.

@@

Perhaps the greatest challenge for ecology in the 21st century is to forecast the impacts of environmental change [@Clark2001; @Petchey2015]. Forecasts require sophisticated modeling approaches that fully account for uncertainty and variability in both ecological process and model parameters [@Luo2011a; but see @Perretti2013a for an argument against modeling the ecological process]. This requires large amounts of data collected over broad spatio-temporal extents, and state-of-the-science modeling techniques cannot overcome data limitations. Population models are a case in point.

@@

The inability of most population models to address landscape-scale problems may explain why land managers and conservation planners have embraced species distribution models (SDMs) [see @Guisan2005 for a review]. SDMs typically rely on easy-to-collect presence/absence data [but see @Clark2014 for new methods] and remotely-sensed environmental covariates that allow researchers to model large spatial extents [e.g., @Maiorano2013]. Thus, it is relatively straightforward to parameterize and project SDMs over landscapes and regions, the scales at which many land-use decisions are made. However, the limitations of SDMs are well known [@Pearson2003; @Elith2009; @Araujo2012]. Ideally, researchers would provide managers with landscape-scale population models, combining the extent of SDMs with information about dynamics and species abundances [@Schurr2012; @Merow2014].

@@

To date, most empirical population modelers have relied on individual-level data, with few attempts to capitalize on population-level measures. An important exception was an effort by @Taylor2004 to model the population growth rate of an invasive species to identify the best strategies for invasion control. They used a "density-structured" model where the state variable is a discrete density state rather than a continuous density measure. Such models do not require individual level demographic data and can adequately describe population dynamics.

@@

Addressing climate change questions with models fit to population-level data is potentially problematic. It is individuals, not populations, that respond to climate variables [@Clark2012]. Ignoring this fact amounts to an "ecological fallacy", where inference about the individual relies on statistical inference on the group [@PIANTADOSI1988]. Population growth (or decline) is the outcome of demographic processes like survival, growth, and recruitment that occur at the level of individual plants. Climate can affect each demographic process in unique, potentially opossing, ways [@Dalgleish2011]. These unique climate responses may be difficult to resolve in statistical models based on population-level data where demographic processes are not identifiable. If important climate drivers are missed because they are averaged over in population-level data, then population models built with such data will make uninformative or unreliable forecasts when confronted with novel climate covariates.

Here, we ask whether statistical and population models based on aggregated, population-level data can predict climate change impacts as well as models based on individual-level data. We used a unique demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. In both models, interannual variation is explained, in part, by climate covariates. We then performed simulations to quantify the sensitivities of species' cover to small perturbations in average precipitation and temperature. Individual level models were clearly better, but, unfortunately, still too uncertain to inform management decisions.

@@

An important, but unexpected, result of our analysis was the great uncertainty in forecasts, even for our best model. The typical approach in ecology is to use point estimates of model parameters to project populations forward according to the specified model, usually allowing for some variability around the determinstic process [e.g. @Battin2007; @Jenouvrier2009; @Adler2012]. If we follow tradition and calculate the mean response to climate perturbation with only model error and interannual variation included, the IPM and the QBM produce opposing forecasts for three of four species (Fig. D1). It would be tempting to interpret this inconsistency as further evidence for the superiority of the IPM. However, if we introduce parameter uncertainty, the forecasts are actually indistinguishable (Fig. 6), though the IPM projections are generally more precise (consistent with our cross-validation results). The real story is that both models produce highly uncertain forecasts. For all species, the 90% quantiles of predicted changes in population size overlapped zero; we cannot even predict whether a change in precipitation or temperature will cause populations to increase or decrease. This result held when we tried perturbing climate by 10% and 20% as well.

@@

How might we improve on this state of affairs? First, forecasts could be improved by matching the spatial scale of predictor variables with the spatial scale of observations. One of the major limitations of the models we fit here is that the climate data are at a much larger scale than the individual level observations of plant size. Climate covariates only vary by year, with no spatial variability within years. Thus, even if we fit models to individual-level data, we are missing the key interaction point between weather and individual plants [@Clark2011b] because all observations share the same climate covariates. Demographic studies should be designed with at least plot-level measurements of climate related variables (e.g., soil moisture). Second, accurately detecting climate signals will take even longer time series. Recent theoretical work on detecting climate signals in noisy data suggests that even advanced approaches to parameter fitting require 20-25 year time series [@Tellerinpress]. Third, ecologists need a stronger commitment to reporting uncertainty. Although most modeling studies explicitly consider model uncertainty, parameter uncertainty is often ignored. In some cases this is because the most convenient statistical methods make it difficult to propogate parameter uncertainty. Yet even Bayesian approaches that allow integration of model fitting and forecasting [@Hobbs2015] are not simple when using modeling approaches like integral projection models that separate the model fitting and simulation stages [@Rees2009]. However, as we have done here, it is still possible to include parameter uncertainty by drawing parameter values from MCMC iterations, taking care to draw all parameters from the same chain and iteration to account for their correlations. Only by being honest about our forecasts can we begin to produce better ones, and forecasts reported without parameter error are disingenuous. In many cases, ignoring parameter error is justifiable when investigating basic demographic mechanisms, but it is not justifiable when forecasting is the goal.

@@

Acknowledgments
---------------
This work was funded by the National Science Foundation through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a CAREER award to PBA (DEB-1054040). We thank the original mappers of the permanent quadrats in Montana and the digitizers in the Adler lab, without whom this work would not have been possible. Informal conversations with Stephen Ellner, Giles Hooker, Robin Snyder, and a series of meetings between the Adler and Weecology labs at USU sharpened our thinking. Brittany Teller and two anonymous reviewers provided comments that improved our manuscript. Compute, storage and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged.

### 2015-08-21 manuscript/manuscript.tex
updates ms with Britta's comments

Forecasting future states of populations has taken on new urgency as the
rate of climate change increases. Traditional plant population models,
typically built using detailed demographic observations from one study
site, can address climate change impacts at one location, but are
difficult to scale up to the landscape and regional scales relevant to
management decisions. One proposal to overcome the limitations of
traditional population models is to build models using population-level
data that is much easier to collect over broad spatial scales than
individual-level data. However, such models ignore the fact that climate
drives population growth through its influence on individual
performance. Here, we test the consequences of aggregating individual
responses when forecasting climate change impacts on four perennial
grass species in a semi-arid Montana grassland. We parameterized two
population models, one based on individual-level data (survival, growth
and recruitment) and one on population-level data from the same dataset
(percent cover of quadrats), and compared their accuracy, precision, and
sensitivity to climate. The individual level model was more accurate and
precise than the aggregate level model when predicting out-of-sample
observations. When comparing climate effects from both models, the
population-level model tends to ``miss'' important climate effects from
at least one vital rate for each species. Increasing the sample size at
the population-level would not necessarily reduce forecast uncertainty;
the way to reduce uncertainty is to capture unique climate dependence of
individual vital rates. Our analysis indicates that there is no short
cut to forecasting climate change impacts on plant populations ---
detailed demographic data is essential. Despite the superiority of the
individual-level model, the forecasts it generated still were too
uncertain to be useful for decision-makers. We need new methods to
collect demographic data more efficiently across environmental gradients
in space and time.

@@

Perhaps the greatest challenge for ecology in the 21st century is to
forecast the impacts of environmental change (Clark et al. 2001, Petchey
et al. 2015). Forecasts require sophisticated modeling approaches that
fully account for uncertainty and variability in both ecological process
and model parameters (Luo et al. 2011). This requires large amounts of
data collected over broad spatio-temporal extents, and
state-of-the-science modeling techniques cannot overcome data
limitations. Population models are a case in point.

@@

The inability of most population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions, the scales at which many land-use decisions are
made. However, the limitations of SDMs are well known (Pearson and
Dawson 2003, Elith and Leathwick 2009, Araújo and Peterson 2012).
Ideally, researchers would provide managers with landscape-scale
population models, combining the extent of SDMs with information about
dynamics and species abundances.

@@

To date, most empirical population modelers have relied on
individual-level data, with few attempts to exploit population-level
data. An important exception was an effort by Taylor and Hastings (2004)
to model the population growth rate of an invasive species to identify
the best strategies for invasion control. They used a
``density-structured'' model where the state variable is a discrete
density state rather than a continuous density measure. Such models do
not require individual level demographic data and can adequately
describe population dynamics.

@@

Addressing climate change questions with models fit to population-level
data is potentially problematic. It is individuals, not populations,
that respond to climate variables (Clark et al. 2012). Ignoring this
fact amounts to an ``ecological fallacy'', where inference about the
individual relies on statistical inference on the group (Piantadosi et
al. 1988). Population growth (or decline) is the outcome of demographic
processes like survival, growth, and recruitment that occur at the level
of individual plants. Climate can affect each demographic process in
unique, potentially opossing, ways (Dalgleish et al. 2011). These unique
climate responses may be difficult to resolve in statistical models
based on population-level data where demographic processes are not
identifiable. If population-level data cannot detect important impacts
of climate drivers on populations, then population models built with
such data will make poor forecasts.

Here, we ask whether statistical and population models based on
aggregated, population-level data can predict climate change impacts as
wells as models based on individual-level data. We used a unique
demographic dataset that tracks the fates of individual plants from four
species over 14 years to build two kinds of single-species population
models, traditional models using individual growth, survival, and
recruitment data and alternative models based on basal cover. In both
models, interannual variation is explained, in part, by climate
covariates. We then performed simulations to quantify the sensitivities
of species' cover to small perturbations in average precipitation and
temperature. Individual level models were clearly better, but,
unfortunately, still too uncertain to inform management decisions.

@@

We use an environmentally stochastic IPM (Rees and Ellner 2009) that
includes the climate covariates from the vital rate statistical models.
In all simulations we ignore the random year effects so that interannual
variation is driven solely by climate. We fit the random year effects in
the vital rate regressions to avoid over-attributing variation to
climate covariates. Our IPM follows the specification of Chu and Adler
(2015) where the population of species \emph{j} is a density function
$n(u_{j},t)$ giving the density of sized-\emph{u} genets at time
\emph{t}. Genet size is on the natural log scale, so that $n(u_{j},t)du$
is the number of genets whose area (on the arithmetic scale) is between
$e^{u_{j}}$ and $e^{u_{j}+du}$. The density function for any size
\emph{v} at time $t+1$ is

@@

To see if models based on aggregated data are as sensitive to climate
covariates as models based on individual level data. So, with our fitted
and validated models in hand, we ran simulations for each model type
(IPM and QBM) under four climate perturbation scenarios: (1) observed
climate, (2) precipitation increased by 1\%, (3) temperature increased
by 1\%, and (4) precipitation and temperature increased by 1\%. We ran
the simulations for 2,500 time steps, enough to estimate equilibrium
cover after discarding an initial 500 time steps as burn-in. Each
simulation was run under two parameter scenarios: (1) using mean
parameter estimates and (2) using randomly drawn parameters from the
MCMC chain. We use (1) to detect the overall sensitivity of equilibrium
cover to climate, and we use (2) to show the impact of model and
parameter uncertainty on forecast precision.

@@

An important, but unexpected, result of our analysis was the great
uncertainty in forecasts, even for our best model. The typical approach
in ecology is to use point estimates of model parameters to project
populations forward according to the specified model, usually allowing
for some variability around the determinstic process (e.g. Battin et al.
2007, Jenouvrier et al. 2009, Adler et al. 2012). If we follow tradition
and calculate the mean response to climate perturbation with only model
error and interannual variation included, the IPM and the QBM produce
opposing forecasts for three of four species (Fig. D1). It would be
tempting to interpret this inconsistency as further evidence for the
superiority of the IPM. However, if we introduce parameter uncertainty,
the forecasts are actually indistinguishable (Fig. 6), though the IPM
projections are generally more precise (consistent with our
cross-validation results). The real story is that both models produce
highly uncertain forecasts. For all species, the 90\% quantiles of
predicted changes in population size overlapped zero; we cannot even
predict whether a change in precipitation or temperature will cause
populations to increase or decrease.
\todo{Add in sentence on larger perturbations.}

@@

How might we improve on this state of affairs? First, forecasts could be
improved by matching the spatial scale of predictor variables with the
spatial scale of observations. One of the major limitations of the
models we fit here is that the climate data are at a much larger scale
than the individual level observations of plant size. Climate covariates
only vary by year, with no spatial variability within years. Thus, even
if we fit models to individual-level data, we are missing the key
interaction point between weather and individual plants (Clark et al.
2011b) because all observations share the same climate covariates.
Demographic studies should be designed with at least plot-level
measurements of climate related variables (e.g., soil moisture). Second,
accurately detecting climate signals will take even longer time series.
Recent theoretical work on detecting climate signals in noisy data
suggests that even advanced approaches to parameter fitting require
20-25 year time series (Teller et al., in review). Third, ecologists
need a stronger commitment to reporting uncertainty. Although most
modeling studies explicitly consider model uncertainty, parameter
uncertainty is often ignored. In some cases this is because the most
convenient statistical methods make it difficult to propogate parameter
uncertainty. Yet even Bayesian approaches that allow integration of
model fitting and forecasting (Hobbs and Hooten 2015) are not simple
when using modeling approaches like integral projection models that
separate the model fitting and simulation stages (Rees and Ellner 2009).
However, as we have done here, it is still possible to include parameter
uncertainty by drawing parameter values from MCMC iterations, taking
care to draw all parameters from the same chain and iteration to account
for their correlations. Only by being honest about our forecasts can we
begin to produce better ones, and forecasts reported without parameter
error are disingenuous. In many cases, ignoring parameter error is
justifiable when investigating basic demographic mechanisms, but it is
not justifiable when forecasting is the goal.

@@

This work was funded by the National Science Foundation through a
Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
CAREER award to PBA (DEB-1054040). We thank the original mappers of the
permanent quadrats in Montana and the digitizers in the Adler lab,
without whom this work would not have been possible. Informal
conversations with Stephen Ellner, Giles Hooker, Robin Snyder, and a
series of meetings between the Adler and Weecology labs at USU sharpened
our thinking. Compute, storage and other resources from the Division of
Research Computing in the Office of Research and Graduate Studies at
Utah State University are gratefully acknowledged.

@@

@@

@@

Rees, M., and S. P. Ellner. 2009. Integral projection models for
populations in temporally varying environments. Ecological Monographs
79:575--594.

@@

>>>>

Forecasting future states of populations has taken on new urgency as the
rate of climate change increases. Traditional plant population models,
typically built using detailed demographic observations from one study
site, can address climate change impacts at one location, but are
difficult to scale up to the landscape and regional scales relevant to
management decisions. One proposal to overcome the limitations of
traditional population models is to build models using population-level
data that is much easier to collect over broad spatial scales than
individual-level data. However, such models ignore the fact that climate
drives population growth through its influence on individual
performance. Here, we test the consequences of aggregating individual
responses when forecasting climate change impacts on four perennial
grass species in a semi-arid Montana grassland. We parameterized two
population models, one based on individual-level data (survival, growth
and recruitment) and one on population-level data from the same dataset
(percent cover of quadrats), and compared their accuracy, precision, and
sensitivity to climate. The individual level model was more accurate and
precise than the aggregate level model when predicting out-of-sample
observations. When comparing climate effects from both models, the
population-level model tends to miss important climate effects from at
least one vital rate for each species. Increasing the sample size at the
population-level would not necessarily reduce forecast uncertainty; the
way to reduce uncertainty is to capture unique climate dependence of
individual vital rates. Our analysis indicates that there is no shortcut
to forecasting climate change impacts on plant populations --- detailed
demographic data is essential. Despite the superiority of the
individual-level model, the forecasts it generated still were too
uncertain to be useful for decision-makers. We need new methods to
collect demographic data more efficiently across environmental gradients
in space and time.

@@

Perhaps the greatest challenge for ecology in the 21st century is to
forecast the impacts of environmental change (Clark et al. 2001, Petchey
et al. 2015). Forecasts require sophisticated modeling approaches that
fully account for uncertainty and variability in both ecological process
and model parameters (Luo et al. 2011, but see Perretti et al. 2013 for
an argument against modeling the ecological process). This requires
large amounts of data collected over broad spatio-temporal extents, and
state-of-the-science modeling techniques cannot overcome data
limitations. Population models are a case in point.

@@

The inability of most population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions, the scales at which many land-use decisions are
made. However, the limitations of SDMs are well known (Pearson and
Dawson 2003, Elith and Leathwick 2009, Araújo and Peterson 2012).
Ideally, researchers would provide managers with landscape-scale
population models, combining the extent of SDMs with information about
dynamics and species abundances (Schurr et al. 2012, Merow et al. 2014).

@@

To date, most empirical population modelers have relied on
individual-level data, with few attempts to capitalize on
population-level measures. An important exception was an effort by
Taylor and Hastings (2004) to model the population growth rate of an
invasive species to identify the best strategies for invasion control.
They used a ``density-structured'' model where the state variable is a
discrete density state rather than a continuous density measure. Such
models do not require individual level demographic data and can
adequately describe population dynamics.

@@

Addressing climate change questions with models fit to population-level
data is potentially problematic. It is individuals, not populations,
that respond to climate variables (Clark et al. 2012). Ignoring this
fact amounts to an ``ecological fallacy'', where inference about the
individual relies on statistical inference on the group (Piantadosi et
al. 1988). Population growth (or decline) is the outcome of demographic
processes like survival, growth, and recruitment that occur at the level
of individual plants. Climate can affect each demographic process in
unique, potentially opossing, ways (Dalgleish et al. 2011). These unique
climate responses may be difficult to resolve in statistical models
based on population-level data where demographic processes are not
identifiable. If important climate drivers are missed because they are
averaged over in population-level data, then population models built
with such data will make uninformative or unreliable forecasts when
confronted with novel climate covariates.

Here, we ask whether statistical and population models based on
aggregated, population-level data can predict climate change impacts as
well as models based on individual-level data. We used a unique
demographic dataset that tracks the fates of individual plants from four
species over 14 years to build two kinds of single-species population
models, traditional models using individual growth, survival, and
recruitment data and alternative models based on basal cover. In both
models, interannual variation is explained, in part, by climate
covariates. We then performed simulations to quantify the sensitivities
of species' cover to small perturbations in average precipitation and
temperature. Individual level models were clearly better, but,
unfortunately, still too uncertain to inform management decisions.

@@

We use an environmentally stochastic IPM (Mark and Ellner 2009) that
includes the climate covariates from the vital rate statistical models.
In all simulations we ignore the random year effects so that interannual
variation is driven solely by climate. We fit the random year effects in
the vital rate regressions to avoid over-attributing variation to
climate covariates. Our IPM follows the specification of Chu and Adler
(2015) where the population of species \emph{j} is a density function
$n(u_{j},t)$ giving the density of sized-\emph{u} genets at time
\emph{t}. Genet size is on the natural log scale, so that $n(u_{j},t)du$
is the number of genets whose area (on the arithmetic scale) is between
$e^{u_{j}}$ and $e^{u_{j}+du}$. The density function for any size
\emph{v} at time $t+1$ is

@@

With our fitted and validated models in hand, we ran simulations for
each model type (IPM and QBM) under four climate perturbation scenarios:
(1) observed climate, (2) precipitation increased by 1\%, (3)
temperature increased by 1\%, and (4) precipitation and temperature
increased by 1\%. We ran the simulations for 2,500 time steps, enough to
estimate equilibrium cover after discarding an initial 500 time steps as
burn-in. Each simulation was run under two parameter scenarios: (1)
using mean parameter estimates and (2) using randomly drawn parameters
from the MCMC chain. We use (1) to detect the overall sensitivity of
equilibrium cover to climate, and we use (2) to show the impact of model
and parameter uncertainty on forecast precision.

@@

An important, but unexpected, result of our analysis was the great
uncertainty in forecasts, even for our best model. The typical approach
in ecology is to use point estimates of model parameters to project
populations forward according to the specified model, usually allowing
for some variability around the determinstic process (e.g. Battin et al.
2007, Jenouvrier et al. 2009, Adler et al. 2012). If we follow tradition
and calculate the mean response to climate perturbation with only model
error and interannual variation included, the IPM and the QBM produce
opposing forecasts for three of four species (Fig. D1). It would be
tempting to interpret this inconsistency as further evidence for the
superiority of the IPM. However, if we introduce parameter uncertainty,
the forecasts are actually indistinguishable (Fig. 6), though the IPM
projections are generally more precise (consistent with our
cross-validation results). The real story is that both models produce
highly uncertain forecasts. For all species, the 90\% quantiles of
predicted changes in population size overlapped zero; we cannot even
predict whether a change in precipitation or temperature will cause
populations to increase or decrease. This result held when we tried
perturbing climate by 10\% and 20\% as well.

@@

How might we improve on this state of affairs? First, forecasts could be
improved by matching the spatial scale of predictor variables with the
spatial scale of observations. One of the major limitations of the
models we fit here is that the climate data are at a much larger scale
than the individual level observations of plant size. Climate covariates
only vary by year, with no spatial variability within years. Thus, even
if we fit models to individual-level data, we are missing the key
interaction point between weather and individual plants (Clark et al.
2011b) because all observations share the same climate covariates.
Demographic studies should be designed with at least plot-level
measurements of climate related variables (e.g., soil moisture). Second,
accurately detecting climate signals will take even longer time series.
Recent theoretical work on detecting climate signals in noisy data
suggests that even advanced approaches to parameter fitting require
20-25 year time series (Teller et al. n.d.). Third, ecologists need a
stronger commitment to reporting uncertainty. Although most modeling
studies explicitly consider model uncertainty, parameter uncertainty is
often ignored. In some cases this is because the most convenient
statistical methods make it difficult to propogate parameter
uncertainty. Yet even Bayesian approaches that allow integration of
model fitting and forecasting (Hobbs and Hooten 2015) are not simple
when using modeling approaches like integral projection models that
separate the model fitting and simulation stages (Mark and Ellner 2009).
However, as we have done here, it is still possible to include parameter
uncertainty by drawing parameter values from MCMC iterations, taking
care to draw all parameters from the same chain and iteration to account
for their correlations. Only by being honest about our forecasts can we
begin to produce better ones, and forecasts reported without parameter
error are disingenuous. In many cases, ignoring parameter error is
justifiable when investigating basic demographic mechanisms, but it is
not justifiable when forecasting is the goal.

@@

This work was funded by the National Science Foundation through a
Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370) and a
CAREER award to PBA (DEB-1054040). We thank the original mappers of the
permanent quadrats in Montana and the digitizers in the Adler lab,
without whom this work would not have been possible. Informal
conversations with Stephen Ellner, Giles Hooker, Robin Snyder, and a
series of meetings between the Adler and Weecology labs at USU sharpened
our thinking. Brittany Teller and two anonymous reviewers provided
comments that improved our manuscript. Compute, storage and other
resources from the Division of Research Computing in the Office of
Research and Graduate Studies at Utah State University are gratefully
acknowledged.

@@

Mark, R., and S. P. Ellner. 2009. Integral projection models for
populations in temporally varying environments. Ecological Monographs
79:575--594.

Merow, C., A. M. Latimer, A. M. Wilson, S. M. McMahon, A. G. Rebelo, and
J. A. Silander. 2014. On using integral projection models to generate
demographically driven predictions of species' distributions:
development and validation using sparse data. Ecography 37:1167--1183.

@@

Perretti, C. T., S. B. Munch, and G. Sugihara. 2013. Model-free
forecasting outperforms the correct mechanistic model for simulated and
experimental data. Proceedings of the National Academy of Sciences of
the United States of America 110:5253--5257.

@@

Schurr, F. M., J. Pagel, J. S. Cabral, J. Groeneveld, O. Bykova,
O'HaraR. B., F. Hartig, W. D. Kissling, H. P. Linder, G. F. Midgley, B.
Schröder, A. Singer, and N. E. Zimmermann. 2012. How to understand
species' niches and range dynamics: A demographic research agenda for
biogeography. Journal of Biogeography 39:2146--2162.

@@

Teller, B. J., P. B. Adler, C. B. Edwards, G. Hooker, R. E. Snyder, and
S. P. Ellner. (n.d.). Linking demography with drivers: climate and
competition. Methods in Ecology and Evolution.
